---
title: "WUP_04"
author: "Ambar Arreola"
date: "Summer 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
thedata<-read.csv(file="GSScsv.csv")
```

```{r eval=FALSE}
thedata <- thedata %>%
mutate(across(where(is.character), as.factor))
```




# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We wanted to find out by looking at the data what sort of data we had .We copied write up 3 to read the dataand change the character variables into factor variables and saw that we had a numerical data which was  WORDSUM and we now have a factor variable called DEGREE. so we deducted that we have a numerical ~ categorical problem.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->





```{r}
ggplot(thedata, aes(x = DEGREE, y = WORDSUM)) +
  geom_boxplot(fill = "skyblue", color = "darkblue", outlier.shape = NA) +
  theme_minimal() +
  labs(x = "DEGREE", y = "WORDSUM") +
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("Fancy Box Plot")
```
LT High school would not overlap with junuir college or graduate.Thus showing that it is very likely to be different schooling. In addition Graduate does not overlap with High school,therefore it is likely to be different.Ultimately, LT Highschool and Graduate are not only different from on another they are both different than another type of schooling, such as High School and Junior College.

### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data= thedata)
```
The average score on the WORDSUM was a little over  10.There is a difference between the average score in Graduate school and the average score in Junior College.Thus further providing that the schooling are different than each other.  

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results
```{r}
model1 <- aov(WORDSUM~DEGREE, data=thedata)
summary.aov(model1)
```

There is a significant difference between the schooling of all 5 degrees.



```{r}
# Perform one-way ANOVA
model <- aov(WORDSUM ~ DEGREE, data = thedata)

# Conduct post-hoc test (Tukey's HSD)
posthoc <- TukeyHSD(model)

# Print the results
print(posthoc)


```
Based on the chart LT high school and Graduate are two degrees that are different from one another. In addition, High school is also different than Graduate. 



```{r}

# Assuming you have the "thedata" data frame with "DEGREE" and "WORDSUM" columns

# Load the emmeans package
library(emmeans)

# Perform one-way ANOVA
result <- aov(WORDSUM ~ DEGREE, data = thedata)

# Perform Scheffe's post hoc test
comparison <- emmeans(result, pairwise ~ DEGREE, adjust = "scheffe")

# Print the comparison table
print(comparison)


```

The chosen procedure of ANOVA followed by Scheffe's post hoc test satisfies the assumptions of independence, normality, and homogeneity of variances. ANOVA is appropriate to test for overall differences among multiple groups, and Scheffe's post hoc test allows for pairwise comparisons while controlling for the family-wise error rate. These procedures are suitable for addressing the research question of comparing the mean "WORDSUM" scores across different degree levels and determining if there are significant differences.



<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion
We tried to use a different test which proved to be more difficult than the tukey test. Based on the results of the ANOVA test, we found significant differences in mean "WORDSUM" scores among the different degree levels (Bachelor, Graduate, High School, Junior College, and LT High School) at a 95% confidence level. The p-value of 0.000236 indicates strong evidence against the null hypothesis, suggesting that the degree level has an effect on "WORDSUM" scores.In conclusion, this analysis highlights the presence of statistically significant differences in mean "WORDSUM" scores among different degree levels. However, further investigation is necessary to determine the underlying factors contributing to these differences and to establish any causal relationship. Understanding the intricacies of this relationship requires a more comprehensive analysis that considers additional variables and potential confounding factors.


<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
